---
title: "Gradient Metrics Homework"
author: "<strong>Tahmidul Islam</strong>" #
date: '<i>`r format(Sys.Date(), "%B %d, %Y")`</i>'
output: 
  html_document:
    theme: sandstone            # Keep this theme - our custom CSS is build upon it
    highlight: tango            # Keep this theme - our custom CSS is build upon it
    keep_md: false              # FALSE if you don't want a MD copy
    code_folding: show          # Makes your code chunks foldable. Delete if you don't want that.
    toc: TRUE                   # FALSE if you don't want a TOC
    toc_depth: 3                # Depth of the TOC
vignette: >
  %\VignetteIndexEntry{Gradient Metric Marker Analysis Homework}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{css, echo = FALSE}
pre {
  max-height: 300px;
  float: left;
  width: 910px;
  overflow-y: auto;
}

pre.r {
  max-height: none;
}
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(knitr)
opts_chunk$set(
	comment = "",
	#fig.width = 12, 
	message = FALSE,
	warning = FALSE,
	tidy.opts = list(
		keep.blank.line = TRUE,
		width.cutoff = 150
		),
	options(width = 150),
	eval = TRUE,
	echo = TRUE,
	fig.height =6,
	fig.width = 8,
	fig.align = "left"
)
# print head
# don't print long-lines
```



# Introduction

This assignment is aimed at replicating a real world marketing research solution where the client wants structure their marketing message in the best possible way to maximize the popularity of their app which helps people sleep better. The objectives of this analysis are:

* **Primary**: Investigate how each attribute (and level) influences the overall likeliness to download using an experiment dataset 

* **Secondary**:  Descriptive analysis of the demographic, psychological and behavioral data of the groups of respondents using a survey dataset

# What Message Should be Sent?

The experiment dataset provides 6 attributes considered in the message testing at the experiment. Let's load the data and have a glimpse at it.

```{r}
# load the R package containing the whole analysis and dataset
library(gradientHW)
# overview of the experiment data
head(experiment_data) %>% kable() %>% kable_styling()
```

Now we explore the possible levels of the attributes.

```{r}
# attributes and levels in experiment data
experiment_data %>%
  select(duration:social_proof) %>%
  map(unique) %>% 
  map(~t(t(.x)))
```


For each combination of these levels, the responders rated how much they would like to download the app. These answers can be found in the `answer' column:

- 1 - Very unlikely 
- 2 - Somewhat unlikely 
- 3 - Somewhat likely 
- 4 - Very likely

Note that, the message testing experiment is a ratings-based conjoint on an ordinal scale. Naturally we approach the problem with the conjoint analysis tool. Conjoint analysis helps us to identify important factors and contribution of levels of those factors when a survey is performed about product preference. We can utilize the excellent R package `Radiant' (https://radiant-rstats.github.io/docs) for this task.

First, we modify our data a little to make it more friendly for the conjoint analysis.

```{r}
experiment_data <- experiment_data %>% 
  mutate_if(is.character,as.factor) %>% #Transforming the character variable to factor variables
  select(-task) #remove the task variable
```


The Radiant package allows us to estimate the `Parts-Worth' (PW) model directly. 

```{r}
# conjoint analysis using Radiant package
conjoint_allrespondents <- conjoint(experiment_data, rvar = "answer",  
                                    evar = names(experiment_data)[2:7]) 
summary(conjoint_allrespondents) 
```
The output gives *part-worths*, *importance weights*, and *regression coefficients*. The part-worths and the regression coefficients give the same information: compared to the reference level (the first level of an attribute, the part-worths are always zero). It is easier to understand the results with the plots. Let's inspect the part-worths (the contribution of the level of the factors).

```{r, 	fig.height = 12,	fig.width = 11}
#plotting part-worths
plot(conjoint_allrespondents,plots = "pw")
```

Using these part-worths it is easy to construct the most desirable combination of messages:

* Duration: "3 months"
* Offer: "improve your sleep sustainably"
* Outcome: "breaking bad habits and creating new routines"
* Price: "$20/month"
* rtb: "a program created just for you" 
* Social proof: "a method that has helped thousands" or "scientific evidence"

In essence, the responders are more likely to download the app if combination of these messages are displayed. We can also find out which attribute has the most effect on the decision of the responders.

```{r}
#plotting importance weights
plot(conjoint_allrespondents,plots = "iw")
```

Clearly the price has the most influence over respondent's will to download the app while the duration has the least effect. The other four attributes have somewhat similar contribution to the responses. 





# Who Are Getting These Messages?

The survey data provides a broad overview of the characteristics and behavior of the respondents. The survey data has several components:

* Demographic
* Psychological 
* Behavioral 

To understand the user group, we take the descriptive statistics route and see who are actually going to receive these messages (assuming we have a representative sample of the receiver population).

First we load the survey data and take a glance.

```{r}
# survey data
head(survey_data)  %>% kable() %>% kable_styling(bootstrap_options = c("striped"), full_width = T, font_size = 11) %>%  scroll_box(height = "300px") 
```
## Demographics

The variables are coded so it will be helpful to look at the variables, the levels and the coding.
```{r}
# subsetting demographic variables
survey_demo <- survey_data %>%
  select(response_id, d_urban:s_problem, d_marital:weights) 
# variables and levels
survey_demo %>%  
  select(-response_id, -weights) %>% 
  map(unique)
```
It always easier to explore the data with plots and charts. From the plots below we can get an overview of the respondents; for example there are almost same number of male and female in the survey, mostly from midwest and southern region. 
```{r, 	fig.height = 8,	fig.width = 11}
# barplot of the demographic variables
survey_demo %>% select(-response_id, -weights) %>% 
  mutate_all(as_factor) %>% 
  gather() %>% 
  drop_na() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scale = 'free') +
  geom_bar()
```

Age distribution is also balanced with ranging from 18 to 65+ years with slightly more people with age between 31 to 64 years. Most of the respondents are from urban/suburban areas with atleast some high school level education. The respondents are mainly employed or retired from diverse household income groups. Most of them are married.

Interesting fact is learned when they are asked the how often would they have trouble sleeping at night during a typical week (e.g., falling asleep, staying asleep, getting good quality sleep). None of the subject replied "Never" or "Rarely". Most of the respondents replied they experience sleep trouble at least two nights a week. It appears that an app for helping people to sleep will be useful to many.


Many other aspects were interviewed as well. The questions and the responses are summarized here.

## Personal philosophy on Science, Products and Health


```{r}
# subsetting the philosophy questions
survey_philosophy <-
  survey_data %>%
  select(response_id, contains('m1_philosophy'))
# attribtutes and levels
survey_philosophy %>%  
  select(-response_id) %>% 
  map(unique) 
```

```{r}
# plotting philosophy data
survey_philosophy %>% select(-response_id) %>% 
  mutate_all(as_factor) %>% 
  gather() %>% 
  drop_na() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scale = 'free') +
  geom_bar()
```


## Attitudes & Awareness towards statements related to sleep 
```{r}
# subsetting awareness variables
survey_awareness <-
  survey_data %>%
  select(response_id, contains('m2_awareness'))
# attributes and levels
survey_awareness %>%  
  select(-response_id) %>% 
  mutate_all(as_factor) %>%
  label_to_colnames() %>% 
  names()
```

```{r}
# plotting awareness variables
survey_awareness %>% 
  mutate_all(as_factor) %>% 
  pivot_longer(cols = -response_id,names_to = 'app', values_to = 'response') %>% 
  select(-response_id) %>% 
  mutate_all(as_factor) %>% 
  drop_na() %>% 
  count(app) %>% 
  ggplot(aes(x = reorder(app, -n), y = n)) + geom_col()
```


## Sources of trouble with sleeping at night and any actions taken to improve the quality of sleep
```{r}
# subsetting souree variables
survey_source <-
  survey_data %>%
  select(response_id, contains('source'))

survey_source %>%  
  select(-response_id) %>% 
  mutate_all(as_factor) %>%
  label_to_colnames() %>% 
  names()
```

```{r}
# plotting source variables
survey_source %>% 
  mutate_all(as_factor) %>% 
  pivot_longer(cols = -response_id,names_to = 'source', values_to = 'response') %>% 
  select(-response_id) %>% 
  drop_na() %>% 
  count(source) %>% 
  ggplot(aes(x = reorder(source, -n), y = n)) + geom_col()
```

## Behavioral questions

```{r}
# subsetting behavioral questions 
survey_behavior <-
  survey_data %>%
  select(response_id, contains('behavior'))

survey_behavior %>%  
  select(-response_id) %>% 
  mutate_all(as_factor) %>%
  label_to_colnames() %>% 
  names()
```

```{r}
#plotting behavioral data
survey_behavior %>% 
  mutate_all(as_factor) %>% 
  pivot_longer(cols = -response_id,names_to = 'behavior', values_to = 'response') %>% 
  select(-response_id) %>% 
  mutate_all(as_factor) %>% 
  drop_na() %>% 
  count(behavior) %>% 
  ggplot(aes(x = reorder(behavior, -n), y = n)) + geom_col()
```

## Other questions

```{r}
# subsetting other variables
survey_other <-
  survey_data %>%
  select(response_id, interst_cbt:interest_coach)

survey_other %>%  
  select(-response_id) %>% 
  map(unique)
```

```{r}
# plotting other variables
survey_other %>% select(-response_id) %>% 
  mutate_all(as_factor) %>% 
  gather() %>% 
  drop_na() %>% 
  ggplot(aes(value)) +
  facet_wrap(~key, scale = 'free') +
  geom_bar()
```

# Market Segmentation

The survey collected data on almost 100 variables. It is very difficult to understand the characteristics of the potential consumers by looking at all these variables all together. Thus a dimension reduction technique is needed to visualize and capture the spread and influence of these attributes. Principal component analysis (PCA) is a widely used statistical tool for dimension reduction. We employ PCA on the survey data. 

```{r}
# for reproducibility of the clustering
set.seed(2021)

# creating dataset suitable for PCA
pca_data <- survey_data %>% 
  select(-response_id, -weights) %>% 
  mutate(s_region = case_when(
    s_region == 'MIDWEST' ~ 0,
    s_region == 'NORTHEAST' ~ 1,
    s_region == 'SOUTH' ~ 2,
    s_region == 'WEST' ~ 3
  )) %>% 
  mutate(across(everything(), ~replace_na(.x, 0))) %>% 
  mutate_all(as_numeric)

# PCA and plotting PCA eigenvalues
pca <- prcomp(pca_data, center = T)  #principle component analysis
fviz_eig(pca)
```

From the PCA result, it doesn't look promising since the principal components (PC) have very little explanation power, i.e. not much dimension reduction is possible. Nevertheless, let's take look at the first two PC.
```{r}
# PC1 vs PC2 plot
autoplot(pca)
```

A cluster analysis will allow us to segment the subjects in the survey according to their traits and responses to the survey questions. Using PC, it is hard to segment the respondents in nice separable groups. Therefore, we take help from the hierarchical cluster analysis to investigate wether we can divide the respondents based on their answers of the survey questions.

```{r}
# create distance matrix for hierarchical clustering
dist <- dist(pca_data, method = "euclidean")
# perform hierarchical clustering
hc <- hclust(dist, method = "complete")
# plot dendogram
fviz_dend(hc, show_labels = FALSE, main = "Uncut Dendrogram")
```

From the dendogram we can decide to cut the trees at the height of 18 to get some reasonable number of clusters. 

```{r}
# cutting the dendogram at height = 18
cluster_assign <- cutree(hc, h = 18)
table(cluster_assign)
```


Cutting the tree at 18 gives us 6 total clusters but there are few clusters with very few observation. So we reduce the number of clusters to 4.

```{r}
# plot the dendogram with 4 clusters
fviz_dend(hc, 
          k = 4,
          show_labels = FALSE,
          rect = TRUE,
          rect_fill = TRUE,
          main = "Four cluster Dendrogram")
```

Using the selected number of clusters, we proceed to the final cluster analysis using the k-means clustering with k = 4.

```{r}
# perform k-means clustering with k = 4
kCluster <- kmeans(pca_data, 4)
pca_data <- mutate(fortify(pca), col=kCluster$cluster) 

# plot clustering result with PCA
ggplot(pca_data) +  
  geom_point(aes(x=PC1, y=PC2, fill=factor(col)), size=3, col="#7f7f7f", shape=21) 

# attach cluster indicator to the survey data
survey_data <- survey_data %>% mutate(cluster = kCluster$cluster)
```
Even with four clusters, it is visible that there are many overlapping groups of subjects. It is not expected that the cluster analysis will be able to segment the market effectively. However, we continue with these four clusters and check the demographics of these clusters.

```{r}
survey_demo <- survey_data %>%
  select(response_id, d_urban:s_problem, d_marital:weights, cluster) 

# summarize survey data grouped by clusters with barplots
survey_demo %>% select(-weights) %>% 
  mutate_all(as_factor) %>% 
  pivot_longer(cols = -c(response_id,cluster) ,names_to = 'key', 
               values_to = 'value') %>% 
  drop_na() %>% 
  ggplot(aes(value, fill = cluster)) +
  facet_wrap(~key, scale = 'free') +
  geom_bar(position = position_fill())
```

From the demographic distribution of the respondents by their cluster identity it appears that the following factors supply useful distinguishing features to segment the market:

1. The number of infant children
2. Education
3. Employment 
4. Marital status
5. Age 




